---
title: 'Google Data Analytics Capstone : How Can a Wellness Technology Company Play
  It Smart?'
author: "soon kien yuan"
date: "2022-08-18"
output:
  pdf_document: default
  html_document: default
---

# Scenario

I'm a junior data analyst at Bellabeat, a high-tech maker of women's health goods. Bellabeat is a modest, successful startup with potential to grow in the global smart device industry. Bellabeat founders and CCO Urka Sren believes examining smart device fitness data might help the company develop. I have been requested to study smart device data for one of Bellabeat's products to learn how consumers use smart devices. The  insights will shape the company's marketing approach. I will offer your research and recommendations to Bellabeat's management team.

# Introduction

## About the company

Bellabeat was formed by Urka Sren and Sando Mur. Sren used her artistic experience to make technology that encourages women worldwide. Bellabeat empowers women by collecting data on movement, sleep, stress, and reproductive health. Bellabeat was started in 2013 and has swiftly become a tech-driven women's wellness firm. Bellabeat had various offices and products by 2016. Bellabeat items are sold by a growing number of online shops in addition to their website. The corporation invests in radio, billboards, print, and TV, but focuses on internet marketing.

## Products

1.  **Bellabeat app**

    The Bellabeat app provides users with health data related to their activity, sleep, stress, menstrual cycle, and mindfulness habits. This data can help users better understand their current habits and make healthy decisions. The Bellabeat app connects to their line of smart wellness products.

2.  **Leaf**

    Bellabeat's classic wellness tracker can be worn as a bracelet, necklace, or clip. The Leaf tracker connects to the Bellabeat app to track activity, sleep, and stress.

3.  **Time**

    This wellness watch combines the timeless look of a classic timepiece with smart technology to track user activity, sleep, and stress. The Time watch connects to the Bellabeat app to provide you with insights into your daily wellness.

4.  **Spring**

    This is a water bottle that tracks daily water intake using smart technology to ensure that you are appropriately hydrated throughout the day. The Spring bottle connects to the Bellabeat app to track your hydration levels.

5.  **Bellabeat membership**

    Bellabeat also offers a subscription-based membership program for users. Membership gives users 24/7 access to fully personalized guidance on nutrition, activity, sleep, health and beauty, and mindfulness based on their lifestyle and goals.

# Data Analysis Process

## 1.0 PHASE 1 :ASK

First up, the analysts needed to define what the project would look like and what would qualify as a successful result. So, to determine these things, they **asked** effective questions and collaborated with leaders and managers who were interested in the outcome of their people analysis.

-   Ask effective questions
-   Define the problem
-   Use structured thinking
-   Communicate with others
-   What is the problem you are trying to solve?

### 1.1 Identify the business task

The business task will be associated with a non-Bellabeat smart devices and analyze the smart device usage data in order to gain insight into how people are already using their smart devices.Then, using the data, generate recommendations for the Bellabeat marketing strategy team to understand the trends. These recommendations and trend insights will be used to enhance the features, functionality, and service quality of the Bellabeat app. So, the business task can be summed up as follows:

1.  Analyze the non-Bellabeat smart device usage data in order to gain insight to help Bellabeat marketing strategy team to understand the trends.

2.  Utilizing the trends and insights provided by smart device usage data in order to improve the features, functionality, and overall service quality of the Bellabeat app

### 1.2 Consider key stakeholders

**Urška Sršen:** Bellabeat's cofounder and Chief Creative Officer

**Sando Mur:** Mathematician and Bellabeat's cofounder; key member of the Bellabeat executive team

**Bellabeat marketing analytics team:** A team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat's marketing strategy. You joined this team six months ago and have been busy learning about Bellabeat''s mission and business goals --- as well as how you, as a junior data analyst, can help Bellabeat achieve them.

## 2.0 PHASE 2 : PREPARE

• Understand how data is generated and collected

• Identify and use different data formats, types, and structures

• Make sure data is unbiased and credible

• Organize and protect data

### 2.1 Dataset overview

This case study utilises [FitBit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit) that is publicly accessible on **Kaggle.** The dataset was collected based on the responses of thirty individuals who participated in a distributed survey conducted by **Amazon Mechanical Turk** between **December 3 and December 5, 2016**.

The **thirty people** who participated in the survey are qualified **Fitbit users** who gave their approval to the submission of personal tracker data. This data included **minute-by-minute output for tracking of physical activity, heart rate, and sleep.**

FitBit Fitness Tracker Data consists of 18 csv file in long format. The data should then be converted to wide format in order to reduce data dimensionality for data analysis.

### 2.2 Data credibility

**ROCC** will be applied to detect if there are data credibility issues.

1.  **Reliabile**

    It is **less reliable**. The dataset is comprised of 30 Fitbit users, which is insufficient to represent the eight million active users in 2016. This would result in a margin of error of 23.56% with a confidence level of 95%, which is less reliable based on [margin error calculator](https://github.com/soonkienyuan/DataAnalytics_Capstone_case_study2/blob/main/Margin%20of%20Error%20Calculator.xlsx) .

    Central Limit Theorem (CLT) says that 30 is the smallest number of samples that can be used. The dataset is still valid based on CLT.

    However, it depends on the stakes. Larger samples are needed for reliable results.

2.  **Original:**

    It is **not original**. The data set was produced by Amazon Mechanical Turk responders to a distributed survey, which is considered as third party data. It would have been preferable if FitBit had delivered the data directly.

3.  **Comprehensive**:

    It does not cover every comprehensive aspect. The data are insufficient because they are lacking certain information (e.g., sex, age, genetic disease) that would assist in producing a more precise analysis. As a result, they cannot be considered comprehensive.

    And yet again, the information was gathered over a span of two months, which is insufficient. It is preferable to have data covering a period of at least a year.

    Last, How did they chose thirty people at random? Which strategy they implement? Does it come from a sample that was picked at random and from a place that was chosen at random as well?

4.  **Current**

    It is **not current.**The data was collected between December 3 and December 5 of 2016, which is now a total of six years ago.

5.  **Cited:**

    It is **Cited**. The datasets were generated by respondents to a distributed survey via Amazon Mechanical Turk.

### 2.3 How Data organized

#### 2.3.1 Loading Packages

```{r message=FALSE, include=FALSE}
library(tidyverse)
library(skimr)
```

#### 2.3.2 Importing Data Sets

```{r}
dailyActivity_merged <- read.csv("https://raw.githubusercontent.com/soonkienyuan/DataAnalytics_Capstone_case_study2/main/datasets/dailyActivity_merged.csv")

dailyCalories_merged  <- read.csv("https://raw.githubusercontent.com/soonkienyuan/DataAnalytics_Capstone_case_study2/main/datasets/dailyCalories_merged.csv")

dailyIntensities_merged <- read.csv("https://raw.githubusercontent.com/soonkienyuan/DataAnalytics_Capstone_case_study2/main/datasets/dailyIntensities_merged.csv")

dailySteps_merged <- read.csv("https://raw.githubusercontent.com/soonkienyuan/DataAnalytics_Capstone_case_study2/main/datasets/dailySteps_merged.csv")

sleepDay_merged <- read.csv("https://raw.githubusercontent.com/soonkienyuan/DataAnalytics_Capstone_case_study2/main/datasets/sleepDay_merged.csv")

weightLogInfo_merged <- read.csv("https://raw.githubusercontent.com/soonkienyuan/DataAnalytics_Capstone_case_study2/main/datasets/weightLogInfo_merged.csv")
```

### 2.4 Dataset structure

#### 2.4.1 `dailyActivity_merged`

```{r}
# first view of the data
head(dailyActivity_merged)

```

```{r}
#structure of dataset
str(dailyActivity_merged)
```

```{r}
#skimming

skim_without_charts(dailyActivity_merged)
```

```{r}
#how many participant records exist in the datasets?

as.data.frame(table(dailyActivity_merged$Id)) %>% rename(Id =Var1 )

```

```{r}
#How many partcipants in the datasets

length(table(dailyActivity_merged$Id))
#or 
n_distinct(dailyActivity_merged$Id)
```

#### 2.4.2 `dailyCalories_merged`

```{r}
# first view of the data
head(dailyCalories_merged)

```

```{r}
#structure of dataset
str(dailyCalories_merged)
```

```{r}
#skimming

skim_without_charts(dailyCalories_merged)
```

```{r}
#how many participant records exist in the datasets?

as.data.frame(table(dailyCalories_merged$Id)) %>% rename(Id =Var1 )

```

```{r}
#How many partcipants in the datasets

length(table(dailyCalories_merged$Id))
#or 
n_distinct(dailyCalories_merged$Id)
```

#### 2.4.3 `dailyIntensities_merged`

```{r}
# first view of the data
head(dailyIntensities_merged)

```

```{r}
#structure of dataset
str(dailyIntensities_merged)
```

```{r}
#skimming

skim_without_charts(dailyIntensities_merged)
```

```{r}
#how many participant records exist in the datasets?

as.data.frame(table(dailyIntensities_merged$Id)) %>% rename(Id =Var1 )

```

```{r}
#How many partcipants in the datasets

length(table(dailyIntensities_merged$Id))
#or 
n_distinct(dailyIntensities_merged$Id)
```

#### 2.4.4 `dailySteps_merged`

```{r}
# first view of the data
head(dailySteps_merged)

```

```{r}
#structure of dataset
str(dailySteps_merged)
```

```{r}
#skimming

skim_without_charts(dailySteps_merged)
```

```{r}
#how many participant records exist in the datasets?

as.data.frame(table(dailySteps_merged$Id)) %>% rename(Id =Var1 )

```

```{r}
#How many partcipants in the datasets

length(table(dailySteps_merged$Id))
#or 
n_distinct(dailySteps_merged$Id)
```

#### 2.4.5 `sleepDay_merged`

```{r}
# first view of the data
head(sleepDay_merged)

```

```{r}
#structure of dataset
str(sleepDay_merged)
```

```{r}
#skimming

skim_without_charts(sleepDay_merged)
```

```{r}
#how many participant records exist in the datasets?

as.data.frame(table(sleepDay_merged$Id)) %>% rename(Id =Var1 )

```

```{r}
#How many partcipants in the datasets

length(table(sleepDay_merged$Id))
#or 
n_distinct(sleepDay_merged$Id)
```

#### 2.4.6 `weightLogInfo_merged`

```{r}
# first view of the data
head(weightLogInfo_merged)

```

```{r}
#structure of dataset
str(weightLogInfo_merged)
```

```{r}
#skimming

skim_without_charts(weightLogInfo_merged)
```

```{r}
#how many participant records exist in the datasets?

as.data.frame(table(weightLogInfo_merged$Id)) %>% rename(Id =Var1 )

```

```{r}
#How many partcipants in the datasets

length(table(weightLogInfo_merged$Id))
#or 
n_distinct(weightLogInfo_merged$Id)
```

## 3.0 PHASE 3 : PROCESS
Noticed that he dailyActivity_merged datasets are just a composite of dailyCalories_merged dataset, dailyIntensities_merged dataset, and dailySteps_merged dataset.

Hence, dataset will used in this study would be `dailyActivity_merged`, `sleepDay_merged` and `weightLogInfo_merged`

### 3.1 Loading Packages

```{r message=FALSE, include=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
```

### 3.2 Checking for duplicte data

```{r}
dailyActivity_merged %>% get_dupes()
```

```{r}
sleepDay_merged %>% get_dupes()
```

```{r}
weightLogInfo_merged %>% get_dupes()

```

I discovered three duplicate data in `sleepDay_merged` by using the `getdupes()` function in janitor package.

By running `skim without charts` from the `skimr` package,found that `weightLogInfo_merged` fat variable contains 65 NA values.

Hence, I decided to drop fat column in `weightLogInfo_merged` and duplicate data in `sleepDay_merged`

### 3.3 Remove duplicates and NA

```{r}
sleepDay_merged <- sleepDay_merged %>% distinct()
```

```{r}

weightLogInfo_merged <- subset(weightLogInfo_merged, select = -c(Fat))
```

### 3.4 Tranforming data

The time and date are recorded together in the same column of both the sleep and weight databases, which I found interesting. If I do decide to use dates to analyse the data across the three datasets, it will be most helpful to divide them into `Date` and `Time` columns.

```{r}
weightLogInfo_merged <- weightLogInfo_merged %>% separate(Date, into=c("Date", "Time"), sep=" ")
sleepDay_merged <- sleepDay_merged %>% separate(SleepDay, into=c("Date", "Time"), sep=" ")
```

## 4.0 PHASE 4 :Analyze

### 4.1 Loading Packages

```{r message=FALSE, include=FALSE}
library(tidyverse)

library(ggplot2)


library(reshape2) #melt the data to create heat map
```

### 4.2 Quick summary statistics

For the `dailyActivity_merged`

```{r}

dailyActivity_merged %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes,
         Calories) %>%
  summary()
```

For the `sleepDay_merged`

```{r}
sleepDay_merged %>%  
  select(TotalSleepRecords,
  TotalMinutesAsleep,
  TotalTimeInBed) %>%
  summary()
```

For the `weightLogInfo_merged`

```{r}
weightLogInfo_merged %>% 
  select(WeightKg,BMI ) %>% 
  summary()

```

### 4.3 Data Exploration

#### 4.3.1 `dailyActivity_merged`

```{r}
ggplot(data=dailyActivity_merged, aes(x=TotalSteps, y=TotalDistance)) + geom_point() +geom_smooth(method=lm) 
```

```{r}
ggplot(data=dailyActivity_merged, aes(x=TotalSteps, y=SedentaryMinutes)) + geom_point() + geom_smooth(method=lm) 
```

```{r}
ggplot(data=dailyActivity_merged, aes(x=Calories, y=TotalDistance)) + geom_point() + geom_smooth(method=lm) 
```

Trends for dailyActivity_merged

-   The correlation coefficient between `TotalSteps` and `TotalDistance` is extremely high. This means that the more steps the user walks, the farther they travel.

-   The scatter plot between `TotalSteps` and `SedentaryMinutes` allows for the estimation of the following conclusion: the longer the `SedentaryMinutes`, the more likely the `totalsteps` have decreased. This means the longer Sedentary time of user, the more likely the users walk less.

-   In other words, the longer a user spends on Sedentary activities, the less probable it is that they would engage in physical activity.

-   The scatter plot between `Calories` and `TotalDistance` allows for the estimation of the following conclusion: the higher the TotalDistance, the more likely the `Calories` burn have increase. The greater the `TotalDistance`, the greater the likelihood that the Calories burned will increase.

-   In other words, the longer a user walks, the greater the chance that they will have a caloric deficit.


#### 4.3.2 ``sleepDay_merged``

heatmap

```{r}

# corelation matrix
cormat <- sleepDay_merged %>%  
  select(TotalSleepRecords,
         TotalMinutesAsleep,
         TotalTimeInBed) %>%
  cor() %>% round(2)


#reshape the data for heatmap
melted_cormat <- melt(cormat)

#data visualization
ggplot(data=melted_cormat,aes(x = Var1, y = Var2, fill = value))+
  geom_tile()+scale_fill_gradient(high = "green", low = "red")+
  ggtitle("sleepDay_merged dataset")+ 
  theme(plot.title = element_text(size = 20, face = "bold"))+
  geom_text(aes(label = round(value, 3)))+
  scale_fill_continuous(low = "red", high = "lightgreen")
```

Trends for dailyActivity_merged

- The correlation coefficient between ``TotalTimeInBed`` and ``TotalMinutesAsleep`` is very high and positive based on the heat map.

-  other words, when sleep duration increases, so does total bedtime.

#### 4.3.3 ``weightLogInfo_merged``

```{r}
ggplot(data=weightLogInfo_merged, aes(x=WeightKg, y=BMI)) + geom_point() + geom_smooth(method=lm) 
```

Trends for dailyActivity_merged

- Nothing interesting found. 

- BMI is the result measure that links body weight to height.

### 4.4 Data merging

I decide to merge the data between  ``dailyActivity_merged`` and ``sleepDay_merged``.

The ``weightLogInfo_merged`` will not be combined because there are only 8 users in this dataset, while the total number of users in all other datasets is 33.

In contrast to other datasets, which have 33 users contributing information, the ``weightLogInfo_merged`` datasets only have 8 users contributing data, which makes the information irrelevant and unsuitable for usage.

Hence, the weightLogInfo_merged
```{r}

#put all data frames into list
data_list <- list(dailyActivity_merged, sleepDay_merged)

#merge all data frames in list
combined_data<- data_list %>% reduce(full_join, by='Id')

```

```{r}
ncol(combined_data)

n_distinct(combined_data$Id)
```

## 5.0 PRASE 5 : SHARE

### 5.1 Tableau Dashboard

```{r}
write.csv(combined_data,"C:\\Users\\soonk\\Documents\\GitHub\\DataAnalytics_Capstone_case_study2\\datasets\\combined_data.csv", row.names = FALSE)
```

Tableau will be used to finish part of this PHASE. Refer to the Tableau file (share.twbx).

I created a dashbaord to make it easier to spot trends and obtain insights from the data.



### 5.2 Minutes Asleep vs. Sedentary Minutes


```{r}
ggplot(data=combined_data, aes(x=TotalMinutesAsleep, y=SedentaryMinutes)) + 
geom_point(color='darkblue') + geom_smooth() +
  labs(title="Minutes Asleep vs. Sedentary Minutes")
```
**Interpretation**

- Nothing special trends or findings found

### 5.3 Total Minutes Asleep vs. Total Time in Bed

```{r}
ggplot(data=combined_data, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + 
  geom_point()+ labs(title="Total Minutes Asleep vs. Total Time in Bed")+geom_smooth()

```
**Interpretation**

- Total time in bed seems positively related to total time asleep. So, to help Bellabeat users get better sleep and increase overall customer experiance, the Bellabeat should think about implementing a notification system that prompts them to lie down and relax.

### 5.3 Total Steps vs. Calories
```{r}
ggplot(data=combined_data, aes(x=TotalSteps, y=Calories)) + 
  geom_point() + geom_smooth() + labs(title="Total Steps vs. Calories")

```


- The scatter plot between `Calories` and `TotalStep` allows for the estimation of the following conclusion: the higher the Total Step, the more likely the `Calories` burn have increase.

- In other words, the longer a user walks, the greater the chance that they will have a caloric deficit.

- Consequently, a reward system should be considered to motivate users to walk and exercise more. Consider a comprehensive reward system with targeted tasks. For example, if your total steps exceed 5,000, you will receive 500 points. The points earned can be redeemed for rewards such as Bellabeat merchandise or discounts on other Bellabeat products, allowing the company to expand and promote the product line while increasing revenue and attracting potential customers.

## 5.0 PRASE 6 : ACT


